\acknowledgements{
First and foremost, I would like to thank my boss, Steve. From the first time I met with him, I knew this was a lab I wanted to work in. Thank you for our in-depth conversations, for sharing your passion for science, for challenging me and encouraging me when it seemed like everything was going wrong. Mostly, thank you for believing in me as a scientist.

To the other members of the lab, thank you for all the support you have provided, in big ways and small. Enrico and Katerina, you were two of the first people I met in the lab, and it's been great fun sharing this ride with you. Thanks for your advice and commiseration! Josh and Alex, I've really enjoyed getting to know you guys over the last year and sharing your enthusiasm as you've started your projects. I know there will be frustrating times, but I hope you find your Ph.D.s to be as rewarding of a journey as I have. A big thanks is due to Sang for invaluable cloning advice and for injecting my flies (and injecting them again ... and again!).

Bettina, you've provided amazing support (and wonderful rhubarb tarts!) along the way. Thank you for answering every question, no matter how small or silly, and helping me overcome my fear of R! Jelena, I credit you with teaching me to program for the first time and opening my eyes to the world of bioinformatics. I don't think it's an exaggeration to say that your influence has changed the path of my career. I've been lucky to have you as a friend and mentor.

Many other members of the Genetics Department have shared their friendship, advice and encouragement over the last four years. I won't try to name everyone because I will surely leave someone out, but thanks to you all for fond memories of Happy Hour shenanigans, Tuesday evenings in the Flying Pig and many other adventures!

A special thanks to my boyfriend Sam (\textless3), for broadening my world, encouraging me to learn more about programming, lively scientific and philosophical discussions, and your unwavering support.

I owe a big thanks to my parents, Norm and Theresa. I know it's been hard for you to have me so far away, and I appreciate your putting up with my wanderlust and continuing to be so supportive! On a more pragmatic note, thank you for providing me with food, an internet connection, and a place to work and sleep during the last 6 weeks or so of the thesis-writing process. I love you both. Dad, I know you've been busy conducting your own research. I hope this thesis is up to your standards, and I eagerly anticipate the chance to finally compare our results. 
}
